---
title: "Use rgl 3D plots for geospatial data"
author: "Lorenz Beck"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Plot 3D Spatial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  \usepackage[utf8]{inputenc}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  results = "asis",
  collapse = TRUE,
  comment = "#>"
)
options(rmarkdown.html_vignette.check_title = FALSE)
```

```{r setup, echo=FALSE, include=FALSE, results="asis"}
library(rgl)
library(sf)
library(plotsphere)
setupKnitr(autoprint = TRUE)
```

The functionalities of the package `rgl` (desribed in the previous vignette) are exploited to plot geospatial data sets that have ellipsoidal coordinates on an interactive sphere. As the package `sf` switches switches from the 1.0 version on to default calculations on a sphere, visualizations should be also plotted rather on a sphere than on a flat projection. 

## Settings of 3D plots in `rgl`

3D plots in the package `rgl` uses an euclidean three dimensional space with x, y, z axis, and therefore simulates a cartesian geocentric coordinate system. 

## 1. The Earth is a sphere in 3D plots

To understand how we model the Earth in our 3D plot with `rgl` we take a look at the fitting geocentric coordinate system `"+proj=geocent"`

```
GEOGCRS["unknown",
    DATUM["WGS_1984",
        SPHEROID["WGS 84",6378137,298.257223563],
        AUTHORITY["EPSG","6326"]],
    UNIT["metre",1,
        AUTHORITY["EPSG","9001"]],
    AXIS["Geocentric X",OTHER],
    AXIS["Geocentric Y",OTHER],
    AXIS["Geocentric Z",NORTH]]
```

In the description WKT of the geocentric projection we can see that it uses the `WGS84` ellipsoid with a radius of 6378137 m. Additionally the assumption is that this ellipsoid is centered at `POINT(0 0 0)` in the cartesian 3D space. Using this information the Earth is modelled in a sphere as the following: 

```{r sphere}
# plot our spherical globe
open3d()
ps_globe(alpha=0.4)
axes3d(box=FALSE)
axis3d('x', pos = c(NA, 0, 0), labels = FALSE)
axis3d('y', pos = c(0, NA, 0), labels = FALSE)
axis3d('z', pos = c(0, 0, NA), labels = FALSE)
title3d(main="Geocentric Coordinate Systems", xlab = 'X axis', ylab = 'Y axis', zlab = 'Z axis')
```

## 2. Coordinates as points on 3D plots

Therefore ellipsoidal or preojected coordinates must be projected to the cartesian geocentric coordinate system 

````{r wgs84}
# create an ellipsoidal point with lon 60째 North and lat 47째 East in WGS84
point_ne <- st_as_sfc("POINT(60 47)", crs = 4326)
st_as_text(point_ne, pretty=TRUE)
# create a second ellipsoidal point with lon -130째 and lat -50째 in WGS84
point_sw <- st_as_sfc("POINT(-130 -50)", crs = 4326)
st_as_text(point_sw, pretty=TRUE)
```

As the Earth radius is implied in `WGS84` geographical projections the original points only need two angles to be referenced `(lon lat)`.

```{r geocentric}
# project both points to geocentric coordinate system
point_ne_geoc <- st_transform(point_ne, "+proj=geocent")
st_as_text(point_ne_geoc, pretty=TRUE)
point_sw_geoc <- st_transform(point_sw, "+proj=geocent")
st_as_text(point_sw_geoc, pretty=TRUE)
```

After projecting it into `+proj=geocent` the point now have three dimensions for its reference `(x y z)`. 
For more information about the conversions check the chapter "2.2 Ellipsoidal Coordinates" in "Spatial Data Science" by Edzer Pebesma and Roger Bivand:
https://keen-swartz-3146c4.netlify.app/cs.html#ellipsoidal-coordinates 

```{r spherical point}
close3d()
open3d()
# add the points to the 3D plot (with axis)
ps_globe()
# add point_ne_geoc
points3d(x = st_coordinates(point_ne_geoc)[1,'X'], y = st_coordinates(point_ne_geoc)[1,'Y'], z = st_coordinates(point_ne_geoc)[1,'Z'])
# add point_sw_geoc
points3d(x = st_coordinates(point_sw_geoc)[1,'X'], y = st_coordinates(point_sw_geoc)[1,'Y'], z = st_coordinates(point_sw_geoc)[1,'Z'])
```
This is the underlying magic behind all conversions from ellipsoidal or projected CRS to geocentric coordinates. This basic procedure is done in all function of geometric conversion and plottings, as points are the the spatial primitives of all spatial geometries (including lines, polygons).
On the following we show the function examples and some additional hints.

## 3. POINTS

```{r points}
close3d()
open3d()
# create random ellipsoidal coordinates (lon lat)
random_matrix <- matrix(c(runif(30, -90, 90), runif(30, -180, 180)), nrow=10, ncol=2)
# convert the matrix of random points into a simple feature multipoint
random_points <- st_multipoint(random_matrix)
random_points <- st_sfc(random_points, crs = 4326)
# plot the sf multipoint on the sphere
ps_globe()
ps_points(random_points, color='red', alpha=0.8)
```

## 4. LINES

```{r lines}
close3d()
open3d()
# create random ellipsoidal coordinates (lon lat)
random_matrix <- matrix(c(runif(30, -90, 90), runif(30, -180, 180)), nrow=10, ncol=2)
# convert the matrix of random points into a simple feature linestring
random_line <- sf::st_linestring(random_matrix)
random_line <- sf::st_sfc(random_line, crs = 4326)
# plot the sf linestring on the sphere
ps_globe()
ps_lines(random_line, color = 'green')
```

## 5. POLYGONS

```{r polygon}
close3d()
open3d()
# create ellipsoidal coordinates (lon lat) (random points wouldn't work here as the geometry would be probably invalid due to crossings in itself)
poly <- st_polygon(list(rbind(c(0,23), c(3,21), c(15,32), c(26,12), c(25,-33), c(42,-40), c(32,-54), c(10,-23), c(0,-23), c(0,23))))
# convert the matrix of random points into a simple feature polygon
poly <- st_sfc(poly, crs = 4326)
# plot the sf polygon on the sphere
ps_globe(alpha=0.4)
ps_polygons(poly, color = 'darkgreen')
```
