---
title: "Basic management of 3D plot"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{plot3D}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(rmarkdown.html_vignette.check_title = FALSE)
```

```{r setup, echo=FALSE, include=FALSE, results="asis"}
library(rgl)
setupKnitr(autoprint = TRUE)
```

The package `rgl` is a 3D visualization system based on OpenGL. It provides a medium to high level interface for use in R, currently modelled on classic R graphics, with extensions to allow for interaction.

This section is a short introduction on useful information about `rgl` regarding this package `plotsphere`. For more detailed description of `rgl` see the github vignettes bymain developer Duncan Murdoch: https://dmurdoch.github.io/rgl/dev/

While plot 3D is mainly used in this packages to produce 3D scatterplots we will rather plot graphical elements as spheres, points, arcs and triangles using `xzy.coords` into the 3D space.

## Graphical Elements

There are many and specific graphical elements available in the `rgl` package. However, we will limit those to most necessary we need

Function        | Description         | Type of element  | Geospatial geometry
----------------| --------------------|------------------|-------------------
`points3d`:     | adds points         | primitive        | point
`triangles3d`:  | adds triangles      | primitive        | polygon
`spheres3d`:    | adds spheres        | constructed      | The Earth
`arcs3d`:       | adds spherical arcs | constructed      | line

## Operators on the 3D plots

As we only deal with one active 3D plot, these functions are the basic operators to manage it:

* `open3d` to open a new 3D plot window (prints id number)
* `cur3d` get id of currently opened 3D plot (prints id number)
* `close3d` to close current 3D plot window
* `set3d` set another 3D plot window as the current one 
* `ids3d` get the objects on the current 3D plot window

```{r sphere}
# open device for 3D plotting
open3d()
spheres3d(x=0, y=0, z=0, radius = 6378137, color = "white")
```

```{r basic2}
cur3d() # prints index of currently open 3D plots
open3d() # closes previous one and opens a new 3D plot (print index of opened plot)
cur3d() # prints index of currently open 3D plots
close3d() # closes the current 3D plot
```

Additionally there are some extensive functions:

* `scene3d`: Saves the current plot into a variable
* `rglwidget`: For embedding graphical displays in HTML documents 

## Our basic Earth

Within adding graphical elements, general arguments can be used to control the material (https://dmurdoch.github.io/rgl/dev/articles/rgl.html#materials-1).
Addtionally, advanced lightning can be set by the `light3d` function

To model our basic Earth where all the geospatial data will be plotted on, we use a blue sphere. 

```{r lightning}
# apply different lights on the current plot
open3d()
spheres3d(x=0, y=0, z=0, radius = 6378137, color = "royalblue3", specular = "royalblue3", shininess = 25, alpha = 0.7)
````

Additionally, we want to add some latitudes and longitudes for orientation. Therefore, every 30 ° will be added to the sphere. 
